---
title: "a3_task3_Craig_Stuart"
author: "Craig Stuart"
date: "11/1/2020"
output: html_document
---

```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load necessary packages
library(tidyverse)
library(here)
library(janitor)
library(knitr)
library(lubridate)
```

## Task 3 Objective:
- Use version control, project management, and data wrangling, analysis and visualization skills to explore red and purple urchin samples recorded by Santa Barbara Coastal Long Term Ecological Research program divers. This task will be broken up into multiple sections in which I will compare the distribution of red and purple urchin sizes and then explore the number of urchins samples by site and year. 

#### Read in urchins.csv data
```{r}
sb_urchins <- read_csv(here("data", "urchins.csv"))
```

#### Clean variable names in data 
```{r}
sb_clean_urchins <- sb_urchins %>% 
  clean_names()
```

## Section 1: Size comparisons of purple and red urchins
### 1. Compare the distribution of purple and red urchin sizes

- use `ggplot()` and `geom_histogram` to Visualize the size distribution between purple and red urchins.

```{r}
ggplot(data = sb_clean_urchins, aes(x = size)) +
  geom_histogram(bins = 20) +
  facet_wrap(~common_name) +
  labs(x = "Urchin Size (cm)", y = "Number of Urchins")
```

- use `ggplot()` and `geom_boxplot` to Visualize the size distribution between purple and red urchins in an alternate way.

```{r}
ggplot(data = sb_clean_urchins, 
       aes(x = common_name, y = size)) +
  geom_boxplot() +
  labs(x = "Urchin Type", y = "Urchin Size (cm)")
```

### 2. Create a summary statistics table
- Group urchin color by `common_name` and summarize the mean, median, sd, and sample size(n)
```{r}
urchin_stats <- sb_clean_urchins %>% 
  group_by(common_name) %>% 
  summarise(mean_urchin = mean(size),
            med_urchin = median(size),
            sd_urchin = sd(size),
            n = n())

kable(urchin_stats)
```

```{r}
# Actual difference between urchin population means
6.776489 - 4.469797

# Percent Difference between urchin population means
(6.776489 - 4.469797) / ((6.776489 + 4.469797)/2) * 100
```

### 3. Overall takeaways of size comparison between two urchin species 

Overall, both samples of urchins show a normal distribution of data based on their histogram visualizations, with purple urchins exhibiting less data spread (sd = 1.18 cm) and a leptokurtic distribution, while the red urchins exhibits slightly more spread (sd = 1.89 cm) and a normal kurtosis distribution curve. The means of the purple and red urchins (4.47 cm and 6.78 cm, respectively) are relatively close in size distribution, only having mean difference of 2.31 cm. To calculate the % difference of the two means, I chose to find the difference between the means and divide by the average of the means to get a 41% difference, suggesting a relatively low difference and a higher chance of being taken from the same population. 


## Section 2: Explore Purple Urchin Counts by Site and Year
### 1. Wrangle data to create a subset of purple urchin data
- Filter purple urchins
- Convert date to class
- Create a new column of year and site name 

```{r}
purple_urchin <- sb_clean_urchins %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date_new = mdy(date)) %>% 
  mutate(year_new = year(date_new)) %>% 
  mutate(site_full = case_when(
    site %in% "NAPL" ~ "Naples Reef",
    site %in% "AQUE" ~ "Arroyo Quemado Reef",
    site %in% "CARP" ~ "Carpinteria Reef", 
    site %in% "MOHK" ~ "Mohawk Reef",
    site %in% "IVEE" ~ "Isla Vista Reef"
  ))

class(purple_urchin)
```

### 2. Make a summary table containing the total number of purple urchins observed by site and year
- Summarize the number of purple urchins per site, per year
- Use the `group_by()` and `count()` functions to orgnaize summary table

```{r}
purple_urchin_stats <- purple_urchin %>% 
  group_by(year_new) %>% 
  count(site_full, year_new)

kable(purple_urchin_stats)
```

- Make exploratory graph to visualize counts of purple urchins at each site

```{r}
ggplot(data = purple_urchin_stats, aes(x = year_new, y = n, fill = site_full)) +
  geom_col() +
  facet_wrap(~site_full)
```

















