---
title: "a3_task4_craig_stuart"
author: "Craig Stuart"
date: "11/2/2020"
output: html_document
---

```{r setup, include=TRUE,  warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Attach packages 
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(knitr)
```

## Task 4 Objective: 
- Use version control, project management, and data wrangling, analysis and visualization skills to explore red and purple urchin samples recorded by Santa Barbara Coastal Long Term Ecological Research program divers. I will compare the mean size of purple urchins between the control and annual treatments within the Naples Reef site in January of 2010. I will use basic exploratory graphs to plot the data comparisons between the two treatments to observe the data distribution. Then I will perform a t-test to interpret and understand the distribution of means between the two groups. 

#### Read in urchins.csv data
```{r}
urchins <- read_csv(here("data", "urchins.csv"))
```

#### Clean variable names in urchin data
```{r} 
clean_urchins <- urchins %>% 
  clean_names()
```

### 1. Wrangle Urchin Data
- Use `filter()` to select observations for purple urchins and the Naples Reef site
- Use `mutate()` to format data as mdy and add new date variables
- Use `filter()` to select observation for january of 2010.

```{r}
treatment_urchins <- clean_urchins %>% 
  filter(common_name == "Purple Urchin", site == "NAPL") %>% 
  mutate(new_date = mdy(date),
         new_month = month(new_date),
         new_year = year(new_date)) %>% 
  filter(new_year == 2010, new_month == 1)
```

### 2. Observe Exploratory Graphs
- Plot data in a histogram graph using `geom_histogram` 

```{r}
ggplot(data = treatment_urchins, aes(x = size)) +
  geom_histogram(bins = 10) +
  facet_wrap(~treatment)
```

- Plot data in a quantile-quantile graph using `geom_qq`

```{r}
ggplot(data = treatment_urchins, aes(sample = size)) +
  geom_qq() +
  facet_wrap(~treatment)
```

- Write 1 - 2 sentences describing your main takeaways about the distribution of urchin sizes for each group.

From the histogram distribution plot, we can observe a general normal distribution curve among the urchin sizes for each treatment group, though it's not totally apparent. There also seems to be a slight positive skew in the distribution curve. We can use the quantile-quantile plot to further assess if our distribution is normal by checking if there is a linear relationship. The linear relationship in the qq plot is easy to see and therefore we can confirm that our samples are normally distributed.  


### 3. Create a basic table of descriptive statistics for both groups (including at least the mean, standard deviation, and sample size)

```{r}
urchin_treatmetn_stats <- treatment_urchins %>% 
  group_by(treatment) %>% 
  summarise(mean_urchin = mean(size),
            med_urchin = median(size),
            sd_urchin = sd(size),
            n = n())

kable(urchin_treatmetn_stats)
```

### 4. Perform a two-sample t-test to compare means between the two treatments
- Use the `pull()` function to create a vector of sizes for each treatment
- Use `t.test()` to perform a two-sample t-test

```{r}
annual_treatment <- treatment_urchins %>% 
  filter(treatment == "ANNUAL") %>% 
  pull(size)

control_treatment <- treatment_urchins %>% 
  filter(treatment == "CONTROL") %>% 
  pull(size)

t.test(x = annual_treatment, y = control_treatment)
```

- Write a comprehensive statement comparing the means, providing meaningful discussion of means difference beyond the p-value. 

The mean urchin size of the annual treatment is 3.92 cm and the mean of the control treatment is 3.73 cm. The difference between the means of the urchin sample treatments is very small, at a 0.19 cm difference, suggesting that the means are fairly close together and making it difficult to determine how different the sample means are from one another, at first glance. The 95% confidence interval of -0.13 cm and 0.52 cm takes into consideration the spread of data to further explains the possible range of differences that the two means could be if multiple samples means were taken again and again. The sd of 0.97 cm and 0.87 cm for annual and control, respectively, suggests that the spread of each data are small and that there is less possible overlap of urchin sizes. Less overlap in data points subsequently decreases the the possibility of drawing from populations with the same mean. The p-value helps us contextualize that possibility. A p-value of 0.25 is relatively low and is saying that there is a lower probability of the two sample means being at least this different if drawn from populations with the same mean. 







