---
title: "a3_task4_craig_stuart"
author: "Craig Stuart"
date: "11/2/2020"
output: html_document
---

```{r setup, include=TRUE,  warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Attach packages 
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(knitr)
```

## Task 4 Objective: 
- Use version control, project management, and data wrangling, analysis and visualization skills to explore red and purple urchin samples recorded by Santa Barbara Coastal Long Term Ecological Research program divers. I will compare the mean size of purple urchins between the control and annual treatments within the Naples Reef site in January of 2010. I will use basic exploratory graphs to plot the data comparisons between the two treatments to observe the data distribution. Then I will perform a t-test to interpret and understand the distribution of means between the two groups. 

#### Read in urchins.csv data
```{r}
urchins <- read_csv(here("data", "urchins.csv"))
```

#### Clean variable names in urchin data
```{r} 
clean_urchins <- urchins %>% 
  clean_names()
```

### 1. Wrangle Urchin Data
- Use `filter()` to select observations for purple urchins and the Naples Reef site
- Use `mutate()` to format data as mdy and add new date variables
- Use `filter()` to select observation for janurary of 2010.

```{r}
treatment_urchins <- clean_urchins %>% 
  filter(common_name == "Purple Urchin", site == "NAPL") %>% 
  mutate(new_date = mdy(date),
         new_month = month(new_date),
         new_year = year(new_date)) %>% 
  filter(new_year == 2010, new_month == 1)
```

### 2. Observe Exploratory Graphs
- Plot data in a histogram graph using `geom_histogram` 

```{r}
ggplot(data = treatment_urchins, aes(x = size)) +
  geom_histogram(bins = 10) +
  facet_wrap(~treatment)
```

- Plot data in a quantile-quantile graph using `geom_qq`

```{r}
ggplot(data = treatment_urchins, aes(sample = size)) +
  geom_qq() +
  facet_wrap(~treatment)
```

- Write 1 - 2 sentences describing your main takeaways about the distribution of urchin sizes for each group.




### 3. Create a basic table of descriptive statistics for both groups (including at least the mean, standard deviation, and sample size)

```{r}
urchin_treatmetn_stats <- treatment_urchins %>% 
  group_by(treatment) %>% 
  summarise(mean_urchin = mean(size),
            med_urchin = median(size),
            sd_urchin = sd(size),
            n = n())

kable(urchin_treatmetn_stats)
```

### 4. Perform a two-sample t-test to compare means between the two treatments
- Use the `pull()` function to create a vector of sizes for each treatment
- Use `t.test()` to perform a two-sample t-test

```{r}
annual_treatment <- treatment_urchins %>% 
  filter(treatment == "ANNUAL") %>% 
  pull(size)

control_treatment <- treatment_urchins %>% 
  filter(treatment == "CONTROL") %>% 
  pull(size)

t.test(x = annual_treatment, y = control_treatment)
```

- Write a comprehensive statement comparing the means, providing meaningful discussion of means difference beyong the p-value. 









